\chapter{Axiomatic Foundations}

\section{The Core Types}

The foundation of Alien Mathematics begins not with sets, but with types. Lean 4's Dependent Type Theory provides the rigorous scaffold upon which non-anthropocentric structures are built.

\begin{definition}[The Alien Structure]
An \textit{Alien Structure} is a generalized algebraic topology over an arbitrary ring, equipped with a functional mapping to the real numbers.
\end{definition}

In Lean 4, this is formalized in \texttt{AlienMathematics.lean} as follows:

\begin{lstlisting}[language=caml, basicstyle=\ttfamily\small, backgroundcolor=\color{lightgray!20}]
structure AlienStructure (R : Type*) [Ring R] where
  carrier : Set R
  functional : R → ℝ
\end{lstlisting}

\section{The Axioms of Alien Transformation}

Human mathematicians often assume continuous symmetries. In the alien framework, we relax these assumptions, leading to pathological transformations.

\begin{axiom}[Pathological Linearity]
There exists a class of transformations that preserve addition but violate scalar multiplication in higher dimensions.
\end{axiom}

\section{The Translation Layer}

As described in the \textit{Bourbaki Translation Layer v2.1}, we must bridge the gap between human readable syntax and the raw AST (Abstract Syntax Tree) manipulation performed by the AI mesh. The translation layer enforces strict semantic boundaries.

The primary mechanism is the \texttt{native\_decide} tactic, which bypasses the standard tactic state and evaluates finite discrete structures directly in the kernel via computational reflection. This allows the AI to inject massive combinatorial exact rational witnesses that are structurally "alien" to human aesthetics but perfectly valid within the type system.
